### 2017-12-15 draft_07.tex
Reasonably complete draft

\documentclass[a4paper, onecolumn]{revtex4-1}

%\documentclass[aps,preprintnumbers,onecolumn,amsmath,amssymb,floatfix,pra]{revtex4}

%\documentclass[aps,pre,twocolumn,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
%\documentclass[reprint, amsmath,amssymb,aps,]{revtex4-1}

\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
% \usepackage{enumerate}

% \usepackage{verbatim}% for \begin{comment} \end{comment}

\usepackage{color}
\usepackage{xcolor}
\usepackage{bm}
\usepackage[]{hyperref}
\hypersetup{
    colorlinks = false,
    linkbordercolor = {red},
    citebordercolor = {green},
}

\usepackage{epstopdf}

\begin{document}

\title{Classical Dynamics of Harmonically Trapped Interacting Particles}

\author{
Zhiyu Dong,$^{1,2}$
R. Moessner,$^2$
and Masudul Haque$^{2,3}$}

\affiliation{
  $^1$ Department of Physics and State Key Laboratory of Surface Physics, Fudan University, Shanghai 200433, China\\ 
  $^2$ Max Planck Institute for the Physics of Complex Systems, N{\"o}thnitzer Str.~38, 01187 Dresden, Germany\\
  $^3$ Department of Theoretical Physics, Maynooth University, Co. Kildare, Ireland
}

\date{\today}

\begin{abstract}

  Motivated by the interest in the dynamics of trapped quantum gases, we study the microcanonical
  dynamics of a trapped gas of classical particles interacting with a finite-range interaction of
  tunable strength.  We examine two questions which have been of interest in quantum dynamics: (1)
  the breathing mode (size oscillation) dynamics of the trapped gas and the dependence of the
  breathing frequency on the interaction strength, and (2) the long-time relaxation and possible
  thermalization of the finite isolated gas.

\end{abstract}

\maketitle

\section{Introduction}

The non-equilibrium dynamics of isolated systems has been the subject of a large volume of work in
recent years.  In particular, in the quantum context, the study of unitary quantum evolution of
many-body and few-body systems has undergone an explosive growth.  This interest was partly fueled
by the availability of experimental systems, e.g., trapped ultracold atomic and ionic systems, where
non-dissipative dynamics can be studied explicitly due to excellent isolation.  There are also
foundational reasons for studying quantum dynamics in isolation: recent activity has led to advances
in the understanding of basic principles connecting quantum dynamics, quantum chaos and the
emergence of statistical mechanics and thermalization (CITE).  For classical many-body systems,
explorations of the connections between microscopic dynamical rules, chaos, ergodicity and
statistical mechanics have a much longer history, dating back to the time of Boltzmann and Poincare
\cite{Dorfman_book_1999, Dumas_book_KAMstory, Boltzmann_legacy_book, EckmannRuelle_RMP85}.

Motivated by questions and issues arising in the field of quantum many-body and few-body dynamics,
in this work we will study the classical dynamics of a collection of interacting particles confined
in a one-dimensional harmonic trap.  The analogs of two questions, studied intensively in the
quantum literature, will be investigated for this classical system: (1) the influence of
interactions on the breathing mode frequency of the trapped gas, and (2) the relaxation and possible
thermalization of the finite system under its own evolution, in the absence of any external baths or
dissipation mechanisms.

Experiments with cold atoms, which have an essential motivational role for the current interest in
many-body quantum dynamics, are generally performed in a trapping potential, either optical or
magnetic.  As a result, in the last two decades many questions of many-body quantum physics have
been reformulated in the presence of a harmonic trap.  The presence of a harmonic trap often results
in new dynamical phenomena.  For example, the motion of vortices in a trapped condensate (CITE) has
new features, such as precession, compared to vortex motion in uniform condensates (CITE).  Trapping
also results in new collective phenomena which have no analog in uniform situations: for example,
collective modes like dipole, breathing and quadrupole modes are specific to trapped many-body
systems.  Dipole modes (center of mass oscillations) and breathing modes (size oscillations) appear
even for one-dimensional trapped systems; higher-dimensional systems display additional, more
complicated modes like quadrupole and scissors modes.  Because of the omnipresence of trapping
potentials in cold-atom experiments, dipole and breathing mode dynamics are pervasive across the
field: These modes have been intensively studied and used for diagnostic purposes since the early
days of research with trapped quantum-degenerate gases, and continue to generate interest today
\cite{Naegerl_Science2009, Esslinger_exp_PRL2003, collective_modes_early_papers,
  Tschischik_BHbreathing_PRA_2013, KroenkeSchmelcher_BM, Bonitz_Review, Bouchoule_PRL2014,
  our_superTG_PRA_2015, STG_Astrakharchik_PRL2005, STG_Chen-etal_PRA2010,
  br_mode_long_range_interactions, 1D_breathing_mode_recent, KohnMode, Cominotti_2015,
  ZarembaTowers_PRA2014, Wetterich_JPB2011, fermion2Dbreathing, Stringari_lowDcollective_PRA15,
  Mazets_1Dbreathing_EPJD11}.  (CITE).
%
In one of the best-known experiments on quantum many-body dynamics,
Ref.~\cite{NewtonsCradle_Nature06}, the dynamics of an integrable system (Lieb-Liniger bosons) was
seen to have ultraslow relaxation in a trap, supposedly due to integrability.  This experiment has
led to continuing efforts to understand how a trap breaks integrability (CITE).

The presence of harmonic trapping is thus a paradigmatic aspect in the field of quantum
non-equilibrium physics.  This motivates our study of classical dynamics of trapped interacting
particles.  Similarly motivated by quantum dynamics, classical trapped dynamics of hard rods has
recently been studied in Refs.\ \cite{DoyonSpohn_JSTAT17, Moore_arXiv17.10}, focusing on
integrability-breaking effects of the trap.  In this work, we consider trapped classical particles
which interact via a simple finite-range potential between any pair of particles: a constant
repulsive force acts whenever the particles are less than a certain distance apart.  The interaction
strength can be tuned, leading to different regimes of behavior.

Our first theme will be the breathing mode of the classical interacting gas, and the effect of
interactions on the frequency of the breathing mode.  In the quantum case, this has long been a
topic of interest for both bosons and fermions.  For bosons with short-range (contact) interactions,
the situation has been studied in one, two and three dimensions.  At zero interactions, the
breathing mode is twice the trapping frequency $\omega_0$ in each case, assuming isotropic trapping.
A mean field (Gross-Pitaevskii) calculation gives the frequency to be $\sqrt{D+2}$ times the
trapping frequency, where $D$ is the spatial dimension.  For three-dimensional (3D) traps, the
mean-field description can be expected to be qualitatively correct, thus the breathing mode
frequency is expected to change monotonically from $2\omega_0$ at zero interactions to
$\sqrt{5}\omega_0$ at large interactions.  For 2D traps, the breathing frequency is exactly
$2\omega_0$ for any interaction strength \cite{PitaevskiiRosch_PRA97}, due to a symmetry.  Classical
trapped particles with scale-invariant interaction potentials also have breathing frequency
$2\omega_0$ independent of interaction strength \cite{PitaevskiiRosch_PRA97}.  The interaction
dependence in 1D (trapped Lieb-Liniger gas) is also remarkable: as the interaction is increased, the
breathing frequency first decreases from $2\omega_0$ to the mean-field prediction
$\sqrt{3}\omega_0$, and then at larger interactions increases again, returning to $2\omega_0$ in the
infinite interaction (Tonks-Girardeau) limit \cite{Naegerl_Science2009,
  Tschischik_BHbreathing_PRA_2013, 1D_breathing_mode_recent, Bouchoule_PRL2014,
  KroenkeSchmelcher_BM}.
%
The breathing mode has also been investigated extensively for fermionic systems with contact
interactions \cite{GrimmSmith_unitaryfermions_PRA08, fermion2Dbreathing,
  Stringari_lowDcollective_PRA15}, and also for fermionic and bosonic systems with long-range
interactions \cite{br_mode_long_range_interactions, Bonitz_Review}.  The breathing mode has also
been studied for a high-temperature trapped gas where a Boltzmann-equation description is valid,
both theoretically \cite{Guery-Odelin1999} and experimentally
\cite{Cornell_classsical_3Dexpt_NatPhys15}.

In this work, we use an explicit microscopic model treated microcanonically, rather than a Boltzmann
equation approach.  Unlike the quantum case, the harmonic potential does not impose a length scale
or an energy scale.  The interaction range can thus be scaled away by a redefinition of length,
i.e., by measuring lengths in units of the interaction range.  There are thus only two important
parameters determining the system behavior, namely the interaction strength and the energy per
particle.  We find that the breathing mode frequency has a non-monotonic dependence on the
interaction parameter.  At small interactions, when the particles pass through each other, the
effect of collisions is to slow down the dynamics, so that the breathing frequency drops below the
non-interacting value $2\omega_0$.  At very large interactions, the particles bounce off each other,
a process which speeds up the size oscillations, resulting in the breathing frequency being larger
than $2\omega_0$.

Our second theme is the process of relaxation at longer timescales, and possible thermalization.
This topic is also motivated by intense recent research in the quantum context.  In quantum isolated
systems, thermalization is now understood in terms of the so-called eigenstate thermalization
hypothesis (ETH), which postulates that in ergodic systems the expectation values of observables in
individual eigenstates depend ony on the eigenenergy, and hence represents the thermal value (CITE).
The connection to statistical mechanics is particularly difficult for explicitly finite systems, in
which case ETH is to be understood in terms of finite size scaling (CITE).  In classical systems,
the connection between microscopic dynamics and statistical mechanics has been continually studied
since the 19th century \cite{Dorfman_book_1999, Dumas_book_KAMstory, Boltzmann_legacy_book,
  EckmannRuelle_RMP85}.  It is generally understood that systems with nonzero Lyapunov exponents,
i.e., chaotic or ergodic systems, thermalize in the long-time limit, provided there are enough
degrees of freedom.  As in the quantum case, finite systems are particularly intricate.  In finite
systems, defining thermalization is trickier as most observables will fluctuate or oscillate
substantially.  In addition, the influence of Kolmogorov-Arnold-Moser (KAM) orbits may play a role
in preventing thermalization; it is generally believed that the fraction of phase space where KAM
physics might be relevant decreases rapidly with particle number \cite{Dumas_book_KAMstory}.

In our work, we examine the relaxation of our trapped interacting classical system, explicity for
finite numbers of particles.  We look for thermalization within finite timescales, by sampling
single-particle energies periodically until such timescales and comparing the distribution of
single-particle energies with the Boltzmann distribution.  We show that whether or not the system
thermalizes in this sense depends on the ratio of interaction strength and energy per particle.  We
also show that this result is reflected in the finite-time Lyapunov exponents of the system.
%
We also examine another notion of thermalization, according to which a system is thermalized when it
loses memory of initial state, by following the dynamics of the particle distribution in phase
space.  Remarkably, we find that one can observe energy thermaliztion even at timescales when the
shape of the phase space distribution continues to perform seemingly coherent oscillations.

There is a significant literature on calculating Lyapunov exponents in model statistical-mechanics
systems and thus exploring ergodicity \cite{Dorfman_book_1999} (CITE).  This provides intuition of
what types of microscopic interactions are likely to produce thermalizability, and is an important
step in the program of understanding statistical mechanics ``from the bottom up''.  However, we are
not aware of a significant literature on following explicitly the microcanonical dynamics of
statistical-mechanical or few-body systems.  Several such studies have appeared recently (CITE),
motivated by the quantum dynamics literature, like the present work.  A few earlier studies have
also examined thermalization dynamics in one dimensional gravitational systems \cite{Yawn1997,
  Tsuchiya2000}.
%
Investigating such real-time dynamics can be expected to provide insights on the timescales and
microscopic mechanisms involved in the emergence of statistical mechanics from collections of
particles.

This paper is organized as follows: In Section \ref{sec:preparation}, we introduce the Hamiltonian,
identify and introduce the essential scales, and describe some aspects of the model and our
simulations.  The breathing frequency is treated in Section \ref{sec:breathing}.  We explain the
main features using both real-space and phase-space pictures of the interaction process, obtain
estimates for the interaction- and energy-dependence of the frequency shift, and compare these
predictions with numerics.  Section \ref{sec:relaxation} treats relaxation and thermalization.  We
propose a condition for relaxation in reasonable timescales from considerations of few-body
dynamics, compare the energy distribution with the Boltzmann distribution, show how the largest
finite-time Lyapunov exponent distribution reflects the relaxation condition, and examine shape
dynamics of the particle distribution in phase space.  Concluding remarks appear in Section
\ref{sec:concl}.

%% This paper is organized as follows: In Section \ref{section:preparation}, we perform a scaling to
%% reduce the number of parameters of the system. In addition, we briefly observe the ground state and
%% reveal two dynamically distinct regime of the system, i.e. strong interaction regime and weak
%% interaction regime. In Section \ref{section:breathing frequency}, we evaluate the breathing
%% frequency according to its mechanism, and compare that with our numerical simulation
%% result. Besides, we also introduce a rotating frame method of treating this problem, which is a
%% convenient tool to analyze simple harmonic systems. In section \ref{section:Thermalization}, we
%% focus on the thermalization property. We first examine the canonical distribution of energy to find
%% out the threshold of thermalization, i.e. under what parameter the thermalization time scale is very
%% large. In the end of this section, we measure the Lyapunov Exponent to quantify the time scales and
%% confirm the former discussion. Then we point out that there are two kind of thermalization behavior
%% that could be measured in this system, one is about energy distribution while another is about shape
%% of cloud in phase-space. Since here we find two sense of ``thermalizing", we will redefine our terms
%% -- When a system's energy distribution goes to equilibrium, we still call it ``thermalized".  But
%% when a system's distribution in phase-space goes to equilibrium, we call it ``relaxed" to distiguish
%% it from ``thermalized". So in section \ref{section:Shape} we actually study the ``thermalizing" of
%% the shape of the cloud in phase-space, or the ``relaxation" behavior of the system. There we
%% discovered a long-lasting mode, which turns out to be a quadruple mode in phase-space.

\section{The model and its equilibrium properties}\label{sec:preparation}

\subsection{Model and Scaling}

Our model for the interacting classical gas involves particles with a simple finite-range repulsive
interaction.  Two particles repel each other whenever they are within a distance $\sigma$ from each
other.  The force of repulsion is a constant, $F_0$, within this distance and zero when the distance
is larger. 
%
\begin{equation}
\Big| F(x) \Big| ~=~ \begin{cases}
F_0  &\quad |x|<\sigma\\
0  &\quad |x|>\sigma
\end{cases}
\label{eq:preparation1}
\end{equation}
where $|x|$ is the distance between two particles.  This equation describes the magnitude of the
interparticle force; the direction is always repulsive.

The gas contains $N$ such identical particles, each of mass $m$, in a harmonic trap.  The
Hamiltonian describing the gas is 
\begin{equation}
  H ~=~
  \frac{1}{2}m\omega_0^2\sum_{i}x_i^2
  ~+~ \frac{1}{2}m\sum_i  v_i^2
  ~+~ \sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
%
It is useful to rescale the quantities.  We will measure distance, time, energy and force in units
of $\sigma$, $1/\omega_0$, $m\omega_0^2\sigma^2$ and $m\omega_0^2\sigma$ respectively:
\begin{equation} \label{eq:transform}
  \tilde{x_i} = \frac{x_i}{\sigma}\ , \quad 
  \tilde{H} = \frac{H}{m\omega_0^2\sigma^2}\ ,  \quad \tilde{t} = \omega_0t \ ,  \quad 
\tilde{F_0} = \frac{F_0}{m\omega_0^2\sigma}\ .
\end{equation}
%
Eq.~\ref{eq:preparation2} is then rewritten as
\begin{equation}
  \tilde{H} =
  \frac{1}{2}\sum_{i}\tilde{x_i}^2
  + \frac{1}{2}\sum_i\left(\frac{d\tilde{x}_i}{d\tilde{t}}\right)^2 
  + \sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless
    1}\tilde{F}_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)  
\end{equation}
%
Through this rescaling, we have reduced the number of the parameters in our model to three: the
energy $\tilde{H}$, the interaction strength $\tilde{F_0}$, and the number of particles $N$.  The
rescaling is equivalent to setting $\sigma$, $\omega_0$ and $m$ to $1$; this is what we do in our
numerical simulation.  In the rest of this paper, we will use ``$E$" and ``$F_0$" to denote the
reduced versions $\tilde{H}$ and $\tilde{F_0}$.

We will be concerned with the size of the cloud.  We quantify the size through the root-mean-square
of particle positions $\left\lbrace x_i\right\rbrace$, 
\begin{equation}
R\equiv\left(\overline{x_i^2}\right)^\frac{1}{2} \, ,
\label{eq:def_of_R}
\end{equation}
and refer to this quantity as the `radius' or `cloud radius'.
%
In the dilute gas limit, there are few or no interactions occuring at most times.  Thus the energy
$E$ is dominated by the trap energy, i.e., $E\approx m\omega_0^2\sum_{i}x_i^2= Nm\omega_0^2R^2$, or,
in our units,
\begin{equation}
  E\approx NR^2 \, .
  \label{eq:E_and_R}
\end{equation}
%
This approximation is good when the average distance betwen particles is much larger than the
interaction range $\sigma=1$, i.e., whenever $R\gg1$, which is the regime we consider dynamics in.
Another way of describing this dilute-gas regime is that the gas is far from the ground state
(described in the next subsection).  Neglecting the interaction energy to estimate $E$ is reasonable
even at very large interactions ($F_0\gg E/N$), because for very large interactions, the particles
behave as hard rods which bounce off each other very rapidly, so that at most instants one can
expect no collisions to be taking place, provided the gas is dilute.  It is also expected to be a
good approximation at very small $F_0$ since we can then simply neglect the interaction energy.
Thus, we will use this as a common estimate for $R$ in terms of the energy.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbhp]
\centering
\includegraphics[scale=0.6]{ZhiyuPictures/N=5_GS_pre_2_rev.eps}
\caption{The ground state of a system with $N=5$ particles, shown via the positions of each
  particle, labeled 1 through 5.  At large interactions, the particles position themselves just
  outside the range of interactions of the neighboring particles.}
\label{fig:GS1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

\subsection{Ground states}

The lowest-energy state of the system has zero kinetic energy; in this state the particles find
stationary positions which minimize the trap (potential) and interaction energies.  The trap
potential tries to squeeze the particles towards the trap center, while the interaction tries to push
them apart.

Because of the `Heaviside theta' form of our interaction, at large enough interactions (large $F_0$)
the particles are spaced exactly at distance equal to the range $\sigma$, which is distance $1$ in
our rescaled units.  The interaction is then effectively a `hard-core' interaction.

At very small $F_0$, the particles in the ground state are close enough to the trap center that each
particle interacts with every other particle.  Equating the interaction force due to all the other
particles with the trapping force, one finds that the position of the $i$-th particle is
\begin{equation}
x_i ~=~ \frac{2F_0}{m\omega_0^2} \left( i - \frac{N+1}{2} \right) ~=~ 2F_0 \left( i - \frac{N+1}{2} \right)
\end{equation}
where the particles are labeled $i=1$ to $i=N$ from left to right.  Thus the particles are
equidistant in this regime.  In this `solid'-like state, the low-lying
excitation involves independent oscillation of the particles around their equilibrium position.
%
In this regime, the distance between the leftmost and rightmost particles is at most $\sigma=1$.
Thus, this situation extends up to $F_0 = \frac{1}{2(N-1)}$.  For the $N=5$ case shown in Figure
\ref{fig:GS1}, this behavior is seen up to $F_0 < 1/8$.

In Figure \ref{fig:GS1}, we can see this crossover from the ``solid-like'' limit (left) to the
``hardcore gas'' limit (right).  In between, there is a rich staircase-like structure, as the
particles attempt to minimize the interaction by being at distance $>1$ from as many other particles
as is compatible with the trap energy.

\subsection{Phase space distribution}

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/stationary_and_rotating_frame_t=1-5_5T.pdf}
%
\caption{Upper (lower) panels are snapshots of the cloud configuration in phase space in the
  stationary (rotating) frame, at instants within the first trap period.  Here $N=50$, $F_0=100$,
  and $E=50000$.  }
\label{fig:Breathingfrequency2_0}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/rotating_frame_t=0-80T.pdf}  
\caption{Phase space snapshots of the cloud dynamics viewed in the rotating frame, at longer
  timescales.  The broadening and rotation observed here are interaction effects: in the
  non-interacting gas the particles are stationary in this frame.}
\label{fig:Breathingfrequency2_1}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

A useful way to visualize the state and evolution of the gas is to plot the position and momentum of
each particle, i.e., to plot the locations of the particles in the single-particle phase space.
This will be useful for visualizing both the breathing mode and relaxation.

Figures \ref{fig:Breathingfrequency2_0} and \ref{fig:Breathingfrequency2_1} show such phase space
snapshots.  We generally start with particles distributed at random positions around the trap
center, initially with no velocity (zero momentum).  The initial state thus has the points lined up
along the $X$ axis.  In the absence of interactions, each particle would undergo simple harmonic
oscillation with period $T=2\pi/\omega_0 = 2\pi$.  The point corresponding to each particle executes
clockwise elliptical motion around the $X$-$P$ plane.  (In our units, the elliptical trajectory is
actually circular.)  This results in the initial line remaining a straight line and rotating
clockwise with exact period $T$.  The effect of the interaction is to smear out the line and spread
the points out toward a rotationally invariant distribution in the $X$-$P$ plane.  The top panel of
Figure \ref{fig:Breathingfrequency2_0} shows this during the first period and Figure
\ref{fig:Breathingfrequency2_1} shows this process over a much longer timescale.

Since the rotation of the initial line of points is simple to understand as a single-particle
(non-interacting) effect, we can focus on interaction effects by viewing the phase space in a
`rotating' frame.  The rotating frame is used in the lower panel of Figure
\ref{fig:Breathingfrequency2_0} and in Figure \ref{fig:Breathingfrequency2_1}.  In this picture, the
`real' $X$ and $P$ axes are rotating counter-clockwise.  This rotating frame picture may be regarded
as a classical version of the ``interaction picture'' of quantum dynamics.  This picture highlights
effects of interactions because the other effects are already encoded in the frame rotation.

In the absence of interactions, each point (each particle) is stationary in the rotating-frame
phase-space picture.  As seen in lower panel of Figure \ref{fig:Breathingfrequency2_0} and in Figure
\ref{fig:Breathingfrequency2_1}, interactions cause a gradual distortion of the line as well as some
degree of rotation.
%
The rotation that is visible in the already rotating $X_{\rm rot}$-$P_{\rm rot}$ frame is the
interaction-induced shift of the breathing frequency from the noninteracting value,
$\omega_B=2\omega_0$.  The distortion of the line toward an eventually rotationally invariant
distribution may be regarded as thermalization or ergodicity.  In the next sections we explore these
two interaction-induced effects.

\subsection{Numerical calculations}

We use the Verlet algorithm (molecular dynamics) to numerically simulate the cloud, using particle
numbers between 5 and 50.  Our force is simple, so that calculating the force at each step is
inexpensive, however, the theta function dependence of the force on particle positions requires the
use of fine timesteps at the beginning and end of each collision process.

The initial state is taken to have particles with zero velocity and random positions; hence the line
distribution in the phase space picture.  This has the advantage that the breathing motion is
prominently visible.  In addition, the question of long-time relaxation has the simple
interpretation of evolving from the line distribution to a circularly symmetric distribution.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbp]
\center
\includegraphics[scale=0.32]{ZhiyuPictures/freq_scanF_differentE_log_2_with_oscillation_demo.png}
\caption{Breathing mode frequency measured at different $E$ and $F_0$, for $N=5$ particles. Inset: a
  demonstration of the oscillation of $E_{trap}$, the total potential energy of particles in the
  trap, Since $E_{trap}\propto \sum_{is}{x_i^2}$, it manifests the oscillation behavior of the
  radius $R(t)$ as defined in Eq.~\ref{eq:def_of_R}.  The breathing frequency is obtained from the
  Fourier transform of $R(t)$.}
\label{fig:Breathingfrequency1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

\section{Breathing Frequency}\label{sec:breathing}

We are interested in the oscillations of the size $R(t)$ of the cloud.  It is convenient that our
finite-size simulations start with a line distribution in phase space.  If we started with a state
whose configuration in phase-space deviates only slightly from circular symmetry, the oscillating
amplitude would be too small to be distinguished from noise.

Without interaction, the breathing mode frequency is exactly 2.  This is visualized readily from the
top panel of \ref{fig:Breathingfrequency2_0}, where $R(t)$ is the extent of the distribution in the
horizontal ($X$) direction.  As the line rotates clockwise with frequency $\omega_0=1$, within each
period the line is twice horizontally aligned (maximum $R$) and twice vertically aligned (minimum
$R$), so that the frequency of $R(t)$ is $\omega_B=2$.  When there is interaction, the frequency
will get shifted: $\omega_B = 2+\delta$.  Numerically, we measure the radius of the cloud $R(t)$ and
get the frequency spectrum of its oscillation behavior by Fourier transform. Then we take the peak
frequency near 2 as the breathing mode frequency.

\subsection{Interaction-dependence}

The frequency measured numerically for different $E$ and $F_0$ is shown in Figure
\ref{fig:Breathingfrequency1} for a system with five particles, plotted as a function of $F_0$.  The
points each correspond to breathing-mode dynamics following from a single initial state; there is no
averaging.  The curves therefore show some noise.  However, two prominent features are clear from
these curves.  At large interactions, $F_0\gg E$, the breathing frequency is larger than $2$, and
saturates around a value which decreases with the system energy $E$.  At small interactions, the
breathing frequency is actually smaller than the non-interacting value $2$.

Below, we will provide a detailed phase-space argument for these behaviors.  However, a simple
real-space picture explains both effects qualitatively as well.  For very large $F_0$, when the
interaction is `hard-core'-like, two particles exchange momentum instantaneously during a collision.
By exchanging the labels of the particles during the collision, this can be interpreted as follows:
particle $A$ carrying momentum $P_A$ jumps by distance $\sigma$ to the right, while particle $B$
carries its momentum $P_B$ and jumps by distance $\sigma$ to the left.  In this manner, every
collision will save a particle some time, $\frac{\sigma}{v}$.  This translates into an increase of
the breathing frequency.  The speed $v$ per particle is on average $\sim\sqrt{E/N}$, and the number
of collision each particle experiences in one period of harmonic oscillation is $\sim N$; hence we
obtain $\delta\sim N^{\frac{3}{2}}E^{-\frac{1}{2}}$.  When $F_0$ is smaller, we can no longer think
of the particles as hard rods; the collisions now take finite time during which the speeds of the
particles are actually slowed down and then sped up again, as they either cross paths or bounce from
each other.  When $F_0$ is small enough, this approaching time will at some critical value of $F_0$
consume the time saved by the finite range of the interaction.  This explains why there is a
small-$F_0$ regime for which the shift $\delta$ is negative.  Figure \ref{fig:Breathingfrequency1}
shows that this critical value of $F_0$, where $\delta$ changes sign and $\omega_B$ crosses $2$,
increases with energy.

\subsection{Estimates using the rotating phase space}

Since we are interested in the deviation $\delta=\omega_B-2$ from the non-interacting breathing
frequency, it is useful to work in the rotating frame, in which a non-interacting cloud would be
stationary (non-rotating).  The rotation frequency of the cloud in this frame is then $\delta$.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[h
]
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/bounce.png}
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/pass.png}
\end{minipage}
}
\caption{Schematic diagrams of two-particle collision process in the $X_{\rm rot}$-$P_{\rm rot}$
  plane. \textbf{(a)}When $F_0$ is large, particles cannot pass each other.  Being subject to a
  constant repulsive force, their trajectories during collision are circular arcs.  For very strong
  interactions (red dashed line, straight), they exchange momenta instantaneously.  The yellow,
  green, and blue dashed curves are for successively weaker interactions.  \textbf{(b)} When $F_0$
  is small, particles pass each other, so that the force they feel is not continuous.  Their
  trajectory has a turning point. The dashed bright blue trajectory is identical with the one in
  (a), which is the critical situation between passing and bouncing. In that case, particles have
  zero relative velocity when they collide. }
\label{fig:Breathingfrequency3}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

To analyze the rotation relative to the $X_{\rm rot}$-$P_{\rm rot}$ frame, we consider two-particle
collisions.  In Figure \ref{fig:Breathingfrequency3}(a), we show schematics of such a collision near
the trap center at strong interactions.  For very large $F_0$, the particles simply exchange
momentum when they are at distance $\sigma$ from each other; this process is shown in red as two
straight lines.  (The particles are initially at $X=\pm\frac{1}{2}\sigma$ and remain at these $X$
values, but exchange their momenta.)  For smaller $F_0$, there is change of both position and
momentum, as shown in yellow, green and blue for successively weaker interactions. The blue line
shows a case where $F_0$ is smaller enough such that the particles can actually cross each other.

In Figure \ref{fig:Breathingfrequency3}(a) we show collisions for small $F_0$, so that the particles
pass each other.  The force changes direction discontinuously when the particles cross.  In the
$X_{\rm rot}$-$P_{\rm rot}$ plane, this is seen as a sharp turning point in the trajectory.

We now estimate $\delta$ at large $F_0$.  The relevant collision process is that shown by the red
lines in Figure \ref{fig:Breathingfrequency3}(a).  The two particles, which are initially on the
$X_{\rm rot}$ axis, are moved to two other points, on the thick gray line, which deviates a small
angle $\alpha$ away from original configuration.  In every period of the harmonic oscillation, each
particle meets each of the other particles twice.  Half of these collisions (about $N$ collisions)
are between particles with large difference in momentum, which is the process shown in
Fig.\ref{fig:Breathingfrequency3}.  (As for the remaining half of the collisions, colliding
particles have smaller difference in their momenta.  In this estimate we ignore the effects of these
collisions, as they clearly contribute far less to the rotation of the particle distribution.)

The precession angle is $\alpha\sim \sigma/P$.  In our rescaled units, the typical momenta of
interacting partilces is of the order $R$.  (In phase space $R$ can be interpreted either as the
extent in real space when the velocities are small, or as the extent in momentum space when the
particles approach $x=0$, i.e. when the distribution is along the $P$ direction.)  Thus, we can
estimate $\alpha\sim \sigma/R$, i.e., $\alpha\sim 1/R$, since our unit of length is $\sigma$.  This
is the rotation per unit collision.  Since there are $\sim N$ collistions per period $T=2\pi$, we
have the estimate for the interaction-induced rotation per unit time as
\begin{equation}
\delta ~=~ 2 \frac{N}{2\pi} \frac{1}{R} ~=~  \frac{1}{\pi}  N^{3/2}E^{-1/2}
\label{eq:breathingfrequency1}
\end{equation}
The factor $2$ accounts for the fact that each rotation of the elongated cloud in phase space
corresponds to two breathing mode periods.  In the last step we have used the estimate $E\sim
NR^2$.

The effect of interaction on breathing frequency is more complicated at smaller $F_0$.  As we have
seen in Figure \ref{fig:Breathingfrequency1}, the effect of a small $F_0$ is to \emph{decrease} the
breathing frequency from its non-interacting value $\omega_B=2$.  This can also be understood using
phase-space pictures of collision processes.  For $F_0$ is not very big, the finite interaction time
needs to be taken into consideration.  During this time, the motion of each particle in phase-space
is moving exactly along the direction of the real $P$ axis at the rate $F_0/m$. Since the real $P$
axis is rotating counter-clockwise, the particle will follow a circular trajectory, e.g., following
the orange or green dashed lines in Fig.~\ref{fig:Breathingfrequency3}(a).  The angle $\alpha$ that
we used above to estimate $\delta$ thus decreases with the increase of interaction time; it is
smaller for the yellow line and even opposite for the green line.  This explains the negative
contribution of interaction to $\delta$, for small enough values of $F_0$.  At even smaller values
of $F_0$, shown in Fig.~\ref{fig:Breathingfrequency3})(b), the particles cross each other.  The
resulting final values are such that the line joining the post-interaction locations of the two
particles has negative $\alpha$, i.e., is tilted clockwise with respect to the $X_{\rm rot}$ axis,
meaning a negative contribution to the breathing frequency.

We can also estimate the critical value of $F_0$ for a given energy (or alternatively the critical
value of $E$ for a given $F_0$) between positive and negative contributions to the breathing
frequency.  Since the $P$ axis rotates with frequency $\omega_0=1$ in the $X_{\rm rot}$-$P_{\rm
  rot}$ plane, the post-collistion position of the particle in the $X_{\rm rot}$-$P_{\rm rot}$ plane
makes angle $\omega_0\tau$ with the red line in Figure \ref{fig:Breathingfrequency3}(a).  Here
$\tau$ is the time over which the interaction acts.  The crossover between positive and negative
$\delta$ is found by comparing this angle to $\alpha$:
\begin{equation}
\omega_0 \tau=\alpha
\end{equation}
The time $\tau$ is approximately the time it takes for the momentum to change sign due to a constant
force $F_0$, since the momentum is $\sim R$, this means $\tau\sim R/F_0$.  Using our previous
estimate  $\alpha\sim \sigma/R$, together with $E\sim NR^2$, gives us the condition
\begin{equation}
  \frac{E}{N} \sim F_0
  \label{eq_freq_turning_condition}
\end{equation}
for the breathing frequency to cross the non-interacting value $\omega_B=2$.  This is roughly the
same criterion for whether two particles will bounce or cross each other in a typical collision.

The condition \eqref{eq_freq_turning_condition} is consistent with Figure
\ref{fig:Breathingfrequency1}, where we noted that the critical $F_0$ grows with increasing energy.

\subsection{Comparisons with numerical data}

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[h]
\centering
\includegraphics[scale=0.6]{ZhiyuPictures/freq_scanF_scanE_pre_Font18.eps}
\caption{Dependence of breathing frequency, calculated for $N=20$ particles, on energy.  Different
  colors represent different $F_0$ --- red for $F_0=2\times10^3$, yellow for $F_0=3\times10^3$,
  green for $F_0=1\times10^4$, and blue for $F_0=1\times10^5$.  Runs with three different initial
  states are shown in each case.  In each run, the frequency is measured both in the time window
  0-100 (labeled as crosses) and in the time window 900-1000 (labeled as circles). The black dashed
  line is the prediction of Eq.~\eqref{eq:breathingfrequency1}.}
\label{fig:Breathingfrequency4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

In Figure \ref{fig:Breathingfrequency4} we show a quantitative test of the main rotating-phase-space
prediction, Eq.~\eqref{eq:breathingfrequency1}.  For large $F_0$, the frequency indeed scales as
$E^{-1/2}$ and for $F_0=10^5$ is quite close to the actual numerical values prediction
\eqref{eq:breathingfrequency1}.  Even for smaller $F_0$, at small enough energies (larger $E^{-1/2}$
values) the breathing mode becomes proportional to  $E^{-1/2}$.

One might ask whether the breathing frequency predictions are also valid at late times, after the
system has `relaxed' from the initial line distribution in phase space to a more spread-out cloud,
as we have seen in Figure \ref{fig:Breathingfrequency2_1}.  In Figure \ref{fig:Breathingfrequency4}
we show the breathing frequency calculated from $R(t)$ oscillations in the first 100 time units, and
also the frequency from data in a later time window. The frequencies appear to be overall stable and
dependent primarily on $F_0$ and $E$.

\section{Relaxation and Thermalization}\label{sec:relaxation}

For any nonzero interaction, the system is expected to be ergodic for $N>2$.  Once there are many
particles, one expects thermalization in the long-time limit.  From the few-particle perspective,
several interesting questions pose themselves.  First, although we expect ergodicity, the question
of how long it takes to thermalize is an open question for small $N$. We treat below a
coarse-grained version of this question: namely, we ask whether particles show ergodic behavior
within reasonable timescales chosen to be (somewhat arbitrarily) in the timestep range of
$10^3$--$10^4$.  Another question is the connection between thermalization, as defined by the
appearance of a Boltzmann distribution and other intuitive characteristics of ergodic behavior such
as whether the single-particle phase space is isotropically occupied.  We find cases where one
aspect is seen while the other is not.

% In the subsections below, we first ***  then ****

\subsection{Few-particle considerations}

Since we are interested in thermalization within finite timescales, it is useful to first consider
mechanisms which hinder relaxation or thermalization.  We begin with few-particle motion.  Since
ergodicity and relaxation are generally expected to be more robust and efficient with larger
particle numbers, consideration of small particle numbers will highlight effects which slow down
relaxation.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbph]
\centering
\includegraphics[width=0.9\textwidth]{ZhiyuPictures/two_particles_a_01.pdf}
\caption{Two-particle dynamics.  Left: Effective potential for relative motion.  Center and right:
  evolution of the individual energies of the two particles.  The two panels correspond to internal
  energy $E_i=2$ and $E_i=5$; the interaction is $F_0=1$.  The beat frequency is lower for larger
  $E_i$. }
\label{fig:thermalization2}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

\subsubsection{Two particles}

We consider the two-particle motion in their center of mass frame.  The center of mass itself
executes simple harmonic oscillation.  In the absence of interactions ($F_0=0$), the effective
potential within the center of mass frame is itself parabolic, with the same frequency.  So their
relative motion is also a harmonic oscillation, with the same frequency $\omega_0$.

With interactions, a term $F_0(1-|r|)\theta(1-|r|)$ is added, where $r$ is the relative
displacement; as shown in Figure \ref{fig:thermalization2}.  Now, the relative motion is no longer
harmonic.  When the internal energy is much larger than $F_0$, one could regard the resulting motion
as having a slightly different frequency, or a collection of frequencies whose center is shifted
slightly from $\omega_0$.  Since the center-of-mass motion is still of frequency $\omega_0$, we have
a superposition of slightly different frequencies, resulting in beating dynamics.  This is clearly
seen in the time evolution of individual energies shown in Figure \ref{fig:thermalization2}.  The
two panels correspond to different internal energy $E_i$ (defined as the total energy minus the
center-of-mass energy).  For larger $E_i$, the distortion of the effective potential (at constant
$F_0$) plays a smaller role in shifting the effective frequency of relative motion; hence the beat
frequency is smaller.

This illustrates a simple mechanism hindering the redistribution of energy between particles, which
is necessary for thermalization or relaxation.  As seen in the example dynamics shown in Figure
\ref{fig:thermalization2}, the difference in energy between the two particles is sustained over
time.  Of course thermalization is not expected anyway in a two-particle system, but we will see
below how this basic effect continues to play a role for larger $N$.

\subsubsection{More particles}

Once we have more than two particles, we expect chaotic or ergodic behavior.  However, it is easy to
imagine mechanisms which drastically slow down redistribution of energy.

Generally, whenever we have some particles with energy very different from others, the dynamics
occurs mostly independently within particle groups of different energies.  For example, with three
particles, consider the situation where two of them, say A and B, have small internal energy, which
means their mutual distance and relative velocity are both small, whereas the third particle (C) has
some energy quite different from A and B.
%
Then, A and B will often be interacting, and if their internal energy is smaller than $F_0$,
significant energy exchange can occur between them.  C will generally exchange little energy with
the pair during interactions, in which (due to larger internal energy between C and either A or B),
the interaction is not effective in energy exchange.
%
An example is shown in Figure \ref{fig:thermalization4}, showing the dynamics of $N=5$ particles.
Two pairs of particles persist in performing `internal' dynamics.  The lone particle participates
only in some slow beating motion with the center of mass of one of the pairs.  Clearly, the energy
mismatch acts as a hurdle to the relaxation process.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{ZhiyuPictures/pair1_pre.eps}
\caption{Time evolution of single-particle energies in a system of five particles, showing a case of
  inefficient energy redistribution.  The energies of the two particles with highest energy (4 \& 5)
  show a clear beating similar to the two-particle case, as does the pair 2 \& 3, although less
  visible.  The `unpaired' particle 1 performs slow beating dynamics with the center of mass of the
  2 \& 3 pair.}
\label{fig:thermalization4}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

%% The argument above still holds in many-particle case. Once we start from some configuration where n
%% particles have a set of close energy levels $\left\lbrace E_n\right\rbrace $ and another m particles
%% have another set of energy levels close to each other $\left\lbrace E_m\right\rbrace $, and
%% $\left\lbrace E_n\right\rbrace $ is quite different from $\left\lbrace E_n\right\rbrace $, we will
%% find these two systems oscillating on the energy level diagram at low frequency.  The discussion
%% above gives us some pictures about the non-ergodic state. For these state, life time is rather long
%% so that once the system reach such configuration (or certain energy distribution), it will take a
%% long time(more than $10^4$ periods for N=5 case) to decay.

\subsection{Relaxation condition and the Boltzmann distribution}

Our consideration of few-particle motion has shown that relaxation is slow when the internal energy
of pairs is large compared to $F_0$.  This allows us to conjecture a condition for relaxation in
reasonable time.  Although it is impossible to express the internal energy of every pair in terms of
the total energy $E$, we can estimate the typical internal energy by the average energy,
$\frac{E}{N}$. At least they are of the same order.  Thus we have the condition
\begin{equation}
F_0 \gtrsim \frac{E}{N}
\label{eq:thermalizatiton_condition}
\end{equation}
for relaxation.  This is the same condition we obtained for the breathing frequency to be increased
(rather than decreased) by the interaction.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbhp]
\centering
%\includegraphics[scale=0.2]{ZhiyuPictures/distribution1_scanF0_E=100_3.eps}
%\caption{Distribution scan $F_0$, $E=100$, $N=5$, $\sigma=1$ {\color{red}{Run again!}} }
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=5_energydistribution_500_400_Font16.eps} 
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=10_energydistribution_500_400_Font16.eps}
% \includegraphics[width=0.32\textwidth]{ZhiyuPictures/N=20_energydistribution_500_400_Font16.eps}
%
\includegraphics[width=0.85\textwidth]{ZhiyuPictures/Boltzmann_2panels_01-crop.pdf}
%
\caption{The distribution of single-particle energies measured in $\sim 10000$ time units. Green
  curves obey the relaxation condition \eqref{eq:thermalizatiton_condition}, and yellow/orange
  corresponds to smaller $F_0$.  In each panel (different $N$), we choose $E/N$ identical ($=1000$),
  so that the threshold interaction between the two behaviors is  $F_0\sim1000$. }
\label{fig:thermalization5}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

Because the model is expected to be ergodic for $N>2$ particles, we expect the same condition to
lead to thermalization.  In fact, the proposed condition works even quantitatively, as we show in
Figure \ref{fig:thermalization5}.  Here, the distribution of energies among the different particles
is presented.  Given our small particle number, a single snapshot of the energies will not yield
enough statistics to investigate the distribution of single-particle energies.  Therefore,
single-particle energies are recorded once every time unit (i.e., at time intervals of $1/\omega_0$)
as the simulation evolves, up to around $t\sim10^4$.  The distribution of these observed values then
shows whether or not the system has thermalized to a Boltzmann distribution.  In the two panels of
Figure \ref{fig:thermalization5}, we show the distributions obtained for a $N=5$ system and a $N=20$
system.  In both cases, we choose the total energy to be $E=1000N$.  In accordance with our
conjectured condition, we find that the observed single-particle distributions are qualitatively
different for $F_0\gtrsim10^3$ and for smaller interactions.  When the ``thermalization condition''
$F_0\gtrsim E/N$ is satisfied, the distribution is roughly exponential ($\sim e^{-\beta\epsilon})$.
For smaller $F_0$, thermalization in this sense is not seen in the single-particle energy
distribution.  Since the systems are expected to be ergodic, the smaller-$F_0$ systems presumably
will also eventually show a Boltzmann distribution of energy, but only at (much) longer timescales.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbph]
\includegraphics[width=0.8\textwidth]{ZhiyuPictures/Boltzmann_log_2panels_01-crop.pdf}
%
\caption{\label{fig:thermalization9} Complaring the single-particle energy distribution
  $\rho(\epsilon)$ with the Boltzmann distribution.  The energy values all satisfy the relaxation
  condition \eqref{eq:thermalizatiton_condition}.  The Boltzmann distribution $\propto
  e^{-\beta\epsilon}$, with the temperature defined as $\beta^{-1}= E/N$, is the thick dashed line.}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

In Figure \ref{fig:thermalization9}, we display the agreement with the Boltzmann distribution in
several `thermalizing' situations, $F_0\gtrsim E/N$.  In each panel, the calculated distributions
from simulations is plotted against the Boltzmann distribution.  The overall agreement shows that
the energy is well-thermalized among the particles on average within the timescales under
consideration.

There are some small deviations from the Boltzmann distribution visible in Figure
\ref{fig:thermalization9}.  Most of the deviation is probably due to numerical noise, and
unsurprising given our small system sizes.  However, some deviation is also expected due to the
contribution of the density of states (DOS).  The probability of finding an energy is actually
proportional to the Boltzmann exponential factor multiplied by the DOS.  For the simple harmonic
oscillator, the DOS is constant.  The interaction deforms the energy shell in the small region
$|x_i-x_j|<\sigma$; so that the DOS is no longer a constant.  The measured distribution seems
systematically slightly smaller than the Boltzmann curve at very small $\epsilon$: when interaction
is weak, particles prefer to stay near zero energy according to the Boltzmann distribution, but when
interaction grows, they are no longer allowed to accumulate near zero energy, in other words the DOS
at low energy decreases with the increase of interaction. Therefore, counts at the low energy part
of Fig.\ref{fig:thermalization9} is slightly lower than the Boltzmann prediction.  The correction to
the density of states is not simple to calculate explicitly.  Since we are in the regime of a dilute
gas ($\sigma\ll R$), the effect is very small.

\subsection{Lyapunov Exponents}

The general idea that thermalization is faster for larger $F_0$ can be seen through the Lyapunov
exponents.  For high-dimensional systems, there is a spectrum of exponents which manifests the
instability of the trajectory along each direction. The largest Lyapunov exponent (LLE) reflects the
shortest time scale for the system lose memory of the initial state.  We follow the method described
in Ref.~\cite{EckmannRuelle_RMP85}.  If the calculation is carried out upto sufficiently late times,
the LLE for an ergodic system is uniquely defined and independent of the starting state.  We would
like to deal with parameter regimes both within and outside our relaxation condition $E<F_0N$ and
$E>F_0N$, and in the latter case we have seen that ergodicity does not become apparent at reasonable
timescales.  Therefore, we carry out the computation of the LLE upto $t\sim10^3$, i.e, we consider
finite-time Lyapunov exponents (CITE).  This means that we can only get
initial-configuration-dependent estimates.  The collection of these estimates forms a distribution
which we show in Figure \ref{fig:LLEdistribution1}.

%% The Lyapunov Exponents is a good tool to quantify the time scale of ``thermalization". Lyapunov
%% Exponents(LE) describes how fast one orbit diverge from its nearby orbits in $\Gamma$-space. But for
%% high dimensional system, LE is a spectrum that manifest the instability of the trajectory along each
%% direction. The largest Lyapunov exponents (LLE) reflects the shortest time scale that system lose
%% its memory. Each point along one trajectory in $\Gamma$-space could have different LLE value, but
%% the distribution of LLE is dependent on macroscopic parameters. As is shown in
%% Fig.\ref{fig:LLEdistribution1}, we measured the LLEs along our trajectory and plot its distribution
%% as a function of total energy.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure}[tbph]
\centering
%\includegraphics[scale=0.4]{ZhiyuPictures/LLEdistribution_1_11_pre_hot1_screenshot.png}
\includegraphics[width=0.55\textwidth]{ZhiyuPictures/LLEdistribution_1_11_pre_hot2_screenshot.png}
\caption{The distribution of largest Lyapunov exponents, for $N=5$, $F_0=1000$, at different
  energies.  The color indicates the observed probability of a measured LLE value being found in a
  bin; the bins are of size 0.01.  }  \label{fig:LLEdistribution1}
\end{figure}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

%% \begin{figure}[h]
%% \centering
%% \includegraphics[scale=0.4]{ZhiyuPictures/LLEdistribution_1_11_pre_bone1_screenshot.png}
%% \includegraphics[scale=0.4]{ZhiyuPictures/LLEdistribution_1_11_pre_bone2_screenshot.png}
%% \caption{``bone". top(bottom):modified(reversed)}
%% \end{figure}

The LLE decreases with the growth of energy. The critical part is near $E = F_0N = 5000$.  Around
this value, the most probable value of LLE decreases well below $\sim1$, which indicates that the
shortest time scale becomes much longer than an oscillation period when $E\gtrsim5000$.  This is
consistent with our predicted thermalization threshold.  As $E$ increases far beyond the
thermalization threshold value, the LLE distribution is close to zero, suggesting a divergence of
relaxation time.
%
We have thus shown an explicit correspondence between energy thermalization in real time and the
(finite-time) LLE.

\subsection{Shape of distribution in single-particle phase space}\label{section:Shape}

Till now, we have studied formulated the thermalization question in terms of energy distributions.
More generally, thermalization may be taken to mean that a many-body system loses memory of its
initial state.  In the single-particle phase-space picture, Figures \ref{fig:Breathingfrequency2_0}
and \ref{fig:Breathingfrequency2_1}, our initial state is very special: The particles are lined up
along the $X$ axis.  The effect of interaction is to distort this line toward a circularly symmetric
distribution.  Thus, we would consider that the memory of the initial state is lost when the
distribution of points in the $X$-$P$ plane are not elongated in any one direction.

To quantify the shape of distribution in phase-space, we define a shape parameter $S$ measuring the
degree of ellipticity:
\begin{equation}
S=\frac{a-b}{a+b}
\end{equation}
where $a$ and $b$ are long axis and short axis of the inertia ellipse in phase space respectively.
These are calculated by diagonalizing the inertia tensor I: $I_{xx}=\sum{p^2},
I_{xp}=I_{px}=-\sum{xp},I_{pp}=\sum{x^2}$.  The shape parameter is $S=1$ for line-shaped
distributions and $S=0$ for circularly symmertic distributions.

%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% 
\begin{figure*}[tbph]
\includegraphics[width=\textwidth]{ZhiyuPictures/combine_shapedynamics_a_01-crop.pdf}
\caption{$N=20, F_0=1000$. Time evolution of the shape parameter, $S(t)$, for $N=20$ particles,
  $F_0=1000$.  The three columns correspond to $E<F_0N$ (low-energy or thermalizing regime),
  $E=F_0N$ (intermediate regime), and $E>F_0N$ (high-energy or non-thermalizing regime).  In the top
  row, coarse-grained values (averaged over 10 periods) are shown, for several initial states with
  the same energy in each case.  The lower row shows original data (not averaged) in a window after
  several thousand time steps.  }
\label{fig:time_evolution_of_S}
\end{figure*}
%%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%% FIGURE %%%%%

The time evolution of the shape parameter $S$ is shown in Fig.\ref{fig:time_evolution_of_S}.
%
The top three pannels show averaged values of $S$ in order to highlight the overall behavior.  Every
data point is the average of the original data over $10T = 20\pi$. The lower six pannels are the
original data measured after several thousand time units.

The coarse-grained (averaged) dynamics on top reflects our intuition that there is slower relaxation
at larger $E/F_0$.  We see a much faster decay for small $E/F_0$, i.e., in the thermalizing regime.
%
The behavior of $S(t)$ at longer times shows (lower panel) that there are long-range coherent
oscillations of the phase-space cloud shape at rather long timescales.  Strikingly, this is even
true for the thermalizing regime (lower left panel), for which the time-averaged single-particle
energies already are Boltzmann-distributed.  Thus, energy thermalization can occur at timescales
much faster than it takes for other features of the gas to relax.

\section{Discussion and Conclusions} \label{sec:concl}

Considering classical interacting particles in a harmonic trap, we have addressed two
non-equilibirum problems motivated by analogous questions in recent research on quantum many-body
dynamics.  We have chosen to work with a simple finite-range interaction with variable interaction
strength, in the `dilute gas' regime where the particles are at average distance significantly
larger than the interaction range.

The first question concerned the breathing mode of the classical trapped gas, and the influence of
interaction on the breathing mode frequency.  We have shown nontrivial dependence of the breathing
frequency on the interaction strength $F_0$: the breathing frequency first decreases as a function
of $F_0$, and around the critical value $F_0 \sim E/N$, in increases above the non-interacting
value.  Eventually at large $F_0$ the frequency shift reaches a plateau at a value
$\propto\sqrt{N^3/E}$.  We have explained this behavior physically, using both real-space and
phase-space pictures.

The second question concerned the relaxation and thermalization behavior of the gas.  This is
particularly interesting for a moderate number of particles, far from the thermodynamic limit.  We
have examined thermalization in the sense of single-particle energy distribution; this has to be
carefully defined when the number of particles is in the `mesoscopic' regime.  Sampling the
single-particle energies from temporal snapshots of the system over a period of time, we find that
the system `thermalizes' in reasonable time for $F_0 \gtrsim E/N$.  We have explained this condition
in terms of relaxation-hindering mechanisms for few-particle systems.  We have also shown that this
threshold corresponds to the largest finite-time Lyapunov exponent exceeding $\approx1$.  Finally,
we have found that (for $F_0 > E/N$) thermalization occurs at timescales much faster than the
timescale for the breathing mode oscillations to damp out.  This is remarkable if one expects
thermalization to correspond to the loss of memory of  initial conditions: the initial shape
ellipticity of the phase space cloud provides very-long-lived oscillations of the shape of the phase
space distribution, even after energy thermalization has already occurred.

The present work opens up various questions.  Our chosen form of interaction is simple and sensible
as a model interaction; and we expect that the $F_0$-dependence of the breathing mode frequency
represents the interaction-dependence for a wide class of interactions.  However, it would be
interesting to ask how collective modes in general depend on the form of interactions.
%
While the connection between ergodicity and thermalization is expected to be simpler in large
systems, for few-body systems like the present case the connection is less clear and raises
questions of timescales.  Our study of a special case exemplifies the need for better general
understanding of the connections between relaxation timescales, interaction strengths,
thermalization, and finite-time Lyapunov expnents.

\begin{thebibliography}{1}

\bibitem{Dorfman_book_1999} J.~P.~Dorfman, ``An Introduction to Chaos in Non-equilibrium Statistical
  Mechanics'', Cambridge University Press, 1999.

\bibitem{Dumas_book_KAMstory} H.~S.~Dumas: ``The KAM Story: A Friendly Introduction to the Content,
History, and Significance of Classical Kolmogorov-Arnold-Moser Theory'', World Scientific, 2014.

\bibitem{Boltzmann_legacy_book} ``Boltzmann's Legacy'', edited by G.~Gallavotti, W.~L.~Reiter, and
  J.~Yngvason; European Mathematical Society (2007).

\bibitem{EckmannRuelle_RMP85}  
  J.-P.~Eckmann and D.~Ruelle,
  ``Ergodic theory of chaos and strange attractors'', 
  \emph{Rev. Mod. Phys.}, {\bf 57}, 617  (1985).

\bibitem{Naegerl_Science2009} E.~Haller, M.~Gustavsson, M.~J.~Mark, J.~G.~Danzl, R.~Hart,
  G.~Pupillo, H.-C.~N\"agerl, ``Realization of an excited, strongly correlated quantum gas phase,''
  {\em Science}, \textbf{325}, 1224 (2009).

\bibitem{collective_modes_early_papers}
S.~Stringari, Phys.\ Rev.\ Lett.\ \textbf{77}, 2360 (1996); 
%
\\ 
%
M.-O.~Mewes, M.~R.~Andrews, N.~J.~van~Druten, D.~M.~Kurn, D.~S.~Durfee, 
C.~G.~Townsend, and W.~Ketterle, Phys.\ Rev.\ Lett.\ \textbf{77}, 988 (1996);
%
\\ 
%
M.~Amoruso, I.~Meccoli, A.~Minguzzi, and M.~P.~Tosi, Eur.\ Phys.\ J.~D \textbf{8}, 361 (2000);
%
\\ 
%
R.~Combescot and X.~Leyronas, Phys.\ Rev.\ Lett.\ \textbf{89}, 190405 (2002);
%
\\ 
%
F.~Chevy, V.~Bretin, P.~Rosenbusch, K.~W.~Madison, and J.~Dalibard, 
Phys.\ Rev.\ Lett.\ \textbf{88}, 250402 (2002);
%
\\ 
%
C.~Menotti and S.~Stringari, ``Collective oscillations of a one-dimensional trapped Bose-Einstein
gas'', {\em Phys.\ Rev.\ A} \textbf{66}, 043610 (2002).
%
\\ 
%
J.~Fuchs, X.~Leyronas, and R.~Combescot, Phys.\ Rev.\ A \textbf{68}, 043610 (2003).

\bibitem{PitaevskiiRosch_PRA97}
%
L.~P.~Pitaevskii and A.~Rosch, {\em Phys.\ Rev.\ A} \textbf{55}, R853 (1997);

\bibitem{Esslinger_exp_PRL2003} H.~Moritz, T.~St\"oferle, M.~K\"ohl, and 
T.~Esslinger,
Phys.\ Rev.\ Lett.\ \textbf{91}, 250402 (2003).

\bibitem{Wetterich_JPB2011}
I.~Boettcher, S.~Floerchinger, and C.~Wetterich, J.~Phys.\ B \textbf{44}, 235301 (2011).

\bibitem{Tschischik_BHbreathing_PRA_2013} W.~Tschischik, R.~Moessner, M.~Haque, Phys.\ Rev.~A 
  \textbf{88}, 063636 (2013).

\bibitem{KroenkeSchmelcher_BM} R.~Schmitz, S.~Kr\"onke, L.~Cao, and 
  P.~Schmelcher, Phys.\ Rev.\ A \textbf{88}, 043601 (2013).

\bibitem{Mazets_1Dbreathing_EPJD11}
I. E. Mazets, ``Integrability breakdown in longitudinaly trapped, one-dimensional bosonic gases'', 
{\em Eur.~Phys.~J.~D} {\bf 65}, 43 (2011).

%%  \bibitem{GaoYu_fermionbreathing_PRA12} 
%% Chao Gao and Zhenhua Yu, 
%%   ``Breathing mode of two-dimensional atomic Fermi gases in harmonic traps'', 
%% {\em Phys. Rev. A} {\bf 86}, 043609 (2012).

\bibitem{fermion2Dbreathing}
%

E.~Vogt, M.~Feld, B.~Fr\"ohlich, D.~Pertot, M.~Koschorreck, and M.~K\"ohl, ``Scale Invariance and
Viscosity of a Two-Dimensional Fermi Gas'', {\em Phys. Rev. Lett.} {\bf 108}, 070404 (2012).
%
\\ 
%
Chao Gao and Zhenhua Yu, ``Breathing mode of two-dimensional atomic Fermi gases in harmonic traps'',
{\em Phys. Rev. A} {\bf 86}, 043609 (2012).
%
\\ 
%
J.~Hofmann, ``Quantum Anomaly, Universal Relations, and Breathing Mode of a Two-Dimensional Fermi
Gas'', {\em Phys. Rev. Lett.} {\bf 108}, 185303 (2012).
%
\\ 
%
S.~K.~Baur, E.~Vogt, M.~K\"ohl, and G.~M.~Bruun, ``Collective modes of a two-dimensional spin-1/2
Fermi gas in a harmonic trap'', {\em Phys. Rev. A} {\bf 87}, 043612 (2013).
%
\\ 
%
B.~P.~van~Zyl, E.~Zaremba, and J.~Towers,
``Collective excitations of a harmonically trapped, two-dimensional, spin-polarized dipolar Fermi
gas in the hydrodynamic regime'', 
{\em Phys. Rev. A} {\bf 90}, 043621 (2014). 
%
\\ 
%
Yi-Cai Zhang and Shizhong Zhang, ``Strongly interacting p -wave Fermi gas in two dimensions:
Universal relations and breathing mode'', {\em Phys. Rev. A} {\bf 95}, 023603 (2017).

%% \bibitem{ZarembaTowers_PRA2014} B.~P.~van~Zyl, E.~Zaremba, and J.~Towers, Phys. Rev. A \textbf{90},
%%   043621 (2014).

\bibitem{Stringari_lowDcollective_PRA15}
G.~De~Rosi and S.~Stringari, ``Collective oscillations of a trapped quantum gas in low dimensions'', 
{\em Phys. Rev. A} {\bf 92}, 053617 (2015).

\bibitem{GrimmSmith_unitaryfermions_PRA08}
S.~Riedl, E.~R.~S\'anchez~Guajardo, C.~Kohstall, A.~Altmeyer, M.~J.~Wright, J.~Hecker~Denschlag,
R.~Grimm, G.~M.~Bruun, and H.~Smith, ``Collective oscillations of a Fermi gas in the unitarity
limit: Temperature effects and the role of pair correlations'', Phys. Rev. A {\bf 78}, 053609
(2008).

\bibitem{Bouchoule_PRL2014}
B.~Fang, G.~Carleo, A.~Johnson, and I.~Bouchoule,  Phys.\ Rev.\ Lett.\ 
\textbf{113}, 035301 (2014).

\bibitem{our_superTG_PRA_2015} W.~Tschischik and M.~Haque, Phys.\ Rev.~A 
  \textbf{91}, 053607 (2015).

\bibitem{Tschischik_ladder_PRA15} W.~Tschischik, R.~Moessner, and M.~Haque, ``Bose-Hubbard ladder
  subject to effective magnetic field: Quench dynamics in a harmonic trap'',
{\em Phys. Rev. A},  {\bf 92}, 023845 (2015).

% Dipole mode of a strongly correlated one-dimensional Bose gas in a split trap: 
% parity effect and barrier renormalization
\bibitem{Cominotti_2015} M.~Cominotti, F.~Hekking, and A.~Minguzzi, 
arXiv:1503.07776.

\bibitem{STG_Chen-etal_PRA2010} S.~Chen, L.~Guan, X.~Yin, Y.~Hao, and X-W.~Guan,   Phys.\ Rev.\ A
  \textbf{81}, 031609(R) (2010).

\bibitem{STG_Astrakharchik_PRL2005} G.~E.~Astrakharchik, J.~Boronat, J.~Casulleras, and
  S.~Giorgini, Phys.\ Rev.\ Lett.\ \textbf{95}, 190407 (2005).

\bibitem{1D_breathing_mode_recent}  
A.~Iu.~Gudyma, G. E.~Astrakharchik, and M.~B.~Zvonarev, ``Reentrant behavior of the
breathing-mode-oscillation frequency in a one-dimensional Bose gas'', 
Phys. Rev. A \textbf{92}, 021601(R) (2015). \\ 
%
S. Choi, V. Dunjko, Z.~D.~Zhang, and M.~Olshanii,
``Monopole Excitations of a Harmonically Trapped One-Dimensional Bose Gas from the Ideal Gas to the
Tonks-Girardeau Regime'', 
{\em Phys. Rev. Lett.} \textbf{115}, 115302 (2015)
%
\\
%
X.-L.~Chen, Y.~Li, and H.~Hu,  Phys.\ Rev.\ A \textbf{91}, 063631 (2015).

\bibitem{br_mode_long_range_interactions} 
%
S.~Bauch, K.~Balzer, C.~Henning, and M.~Bonitz, Phys. Rev. B \textbf{80}, 054515 
(2009);
%
\\
%
J.~W.~Abraham, K.~Balzer, D.~Hochstuhl, and M.~Bonitz, Phys. Rev. B \textbf{86}, 
125112 (2012);
%
\\
%
C.~R.~McDonald, G.~Orlando, J.~W.~Abraham, D.~Hochstuhl, M.~Bonitz, and T.~Brabec, Phys.\ Rev.\
Lett.\ \textbf{111}, 256801 (2013);
%
\\
%
J.~W.~Abraham, M.~Bonitz, C.~McDonald, G.~Orlando, and T.~Brabec, New J.~Phys \textbf{16}, 013001
(2014).

\bibitem{Bonitz_Review} J.~W.~Abraham, and M.~Bonitz, ``Quantum Breathing Mode of Trapped Particles:
  from nanoplasmas to ultracold gases'', {\em Contributions to Plasma Physics}, \textbf{54}, 27
  (2014).

\bibitem{KohnMode}
%
L.~Brey, N.~F.~Johnson, and B.~I.~Halperin, Phys.\ Rev.\ B \textbf{40}, 10647 (1989).
%
\\ 
%
M.~Bonitz, K.~Balzer, and R.~van~Leeuwen, {\em Phys. Rev. B} \textbf{76}, 045341 (2007).

\bibitem{NewtonsCradle_Nature06}  
T.~Kinoshita, T.~Wenger, and D.~S.~Weiss, ``A quantum Newton's cradle'', {\em Nature}, {\bf 440,} 900 (2006).

\bibitem{DoyonSpohn_JSTAT17} B.~Doyon and H.~Spohn, ``Dynamics of hard rods with initial domain wall
  state'', {\em J.~Stat. Mech.}, {\bf 2017}, 073210 (2017).

\bibitem{Moore_arXiv17.10} X.~Cao, V.~B.~Bulchandani, J.~E.~Moore, ``Incomplete thermalization from
  trap-induced integrability breaking: lessons from classical hard rods'', arXiv:1710.09330.

\bibitem{Guery-Odelin1999}
D.~Gu\'ery-Odelin, F.~Zambelli, J.~Dalibard, and S.~Stringari, ``Collective
  oscillations of a classical gas confined in harmonic traps,'' {\em Phys. Rev.
  A}, {\bf 60}, 4851 (1999).

\bibitem{Cornell_classsical_3Dexpt_NatPhys15} D.~S.~Lobser, A.~E.~S.~Barentine, E.~A.~Cornell, and
  H.~J.~Lewandowski, ``Observation of a persistent non-equilibrium state in cold atoms'', Nat Phys
  {\bf 11}, 1009 (2015).

\bibitem{Dalfovo1997}
F.~Dalfovo, S.~Giorgini, M.~Guilleumas, L.~Pitaevskii, and S.~Stringari,
  ``Collective and single-particle excitations of a trapped bose gas,'' {\em
  Phys. Rev. A}, {\bf 56}, 3840 (1997).

\bibitem{Jin1996}
D.~S. Jin, J.~R. Ensher, M.~R. Matthews, C.~E. Wieman, and E.~A. Cornell,
  ``Collective excitations of a bose-einstein condensate in a dilute gas,''
  {\em Phys. Rev. Lett.}, vol.~77, pp.~420--423, July 1996.

\bibitem{Dalfovo1999}
F.~Dalfovo, S.~Giorgini, L.~P. Pitaevskii, and S.~Stringari, ``Theory of
  bose-einstein condensation in trapped gases,'' {\em Rev. Mod. Phys.},
  vol.~71, pp.~463--512, Apr. 1999.

\bibitem{Stringari1996}
S.~Stringari, ``Collective excitations of a trapped bose-condensed gas,'' {\em
  Phys. Rev. Lett.}, vol.~77, pp.~2360--2363, Sept. 1996.

\bibitem{Tsuchiya2000}
T.~Tsuchiya and N.~Gouda, ``Relaxation and Lyapunov time scales in a
  one-dimensional gravitating sheet system,'' {\em Phys. Rev. E}, {\bf 61}, 948 (2000).

\bibitem{Yawn1997}
K.~R. Yawn and B.~N. Miller, ``Ergodic properties and equilibrium of
  one-dimensional self-gravitating systems,'' {\em Phys. Rev. E}, vol.~56,
  pp.~2429--2436, Sept. 1997.

\bibitem{Jin2013}
F.~Jin, T.~Neuhaus, K.~Michielsen, S.~Miyashita, M.~A. Novotny, M.~I.
  Katsnelson, and H.~D. Raedt, ``Equilibration and thermalization of classical
  systems,'' {\em New Journal of Physics}, vol.~15, no.~3, p.~033009, 2013.

\bibitem{GueryOdelin_thermalization_PRA06} M.~Anderlini and D.~Gu\'ery-Odelin, ``Thermalization in
  mixtures of ultracold gases'', {\em Phys. Rev. A} {\bf 73}, 032706 (2006).

\end{thebibliography}

 \end{document}